---
title: "AIRLINE PASSENGERS SATISFACTION: CLASSIFICATION"
output: html_notebook
date: 11/09/2001
---

## Database description
[...]

## Variable description
[...]

# OUTLINE 
1. Data upload
2. Exploratory analysis:
    - check for missing values
    - outliers analysis
    - variables density
    - customer satisfaction distribution for categorical variables
    - customer satisfaction distribution for continuous variables
    - correlation matrix
3. Train and test sets split
4. Logistic regression:
    - general logistic model
    - forward step-wise models analysis
    - cross-validation
    - models comparison based on R-squared, CV error, CP
    - models selection (maybe one model per criteria if they differs)
    - models evaluation (accuracy, specificity, sensitivity, ...)
    - ROC curve
5. Linear Discriminant Analysis:
    - ...
    - compare LDA and logistic
6. Tree predictors:
    - ...
    - comparison with logistic model
7. Conclusion and Interpretation


## Libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(grid)
library(corrr)
```


# 1) DATA
```{r}
df <- read_csv("Database/airline_passenger_satisfaction.csv")
```

```{r}
head(df)
```

### Dataset structure
```{r}
str(df)
```

# 2) EXPLANATORY ANALYSIS

## Check for missing values
```{r}
# Count NA values for each column
na_counts <- sapply(df, function(x) sum(is.na(x)))


# Calculate percentage of NA values for each column
na_percentage <- na_counts / nrow(df) * 100

# Create a data frame with column names, NA counts, and percentages
na_table <- data.frame(Column = names(na_counts), NA_Count = na_counts, Percentage = na_percentage)

# Display the result table
na_table
```

  

