---
title: "AIRLINE PASSENGERS SATISFACTION: CLASSIFICATION"
output: html_notebook
date: 16/06/2023
---

## Database description
The following database contains information about airline customers, the flight they took and the results of a satisfaction questionnaire about their flight experience.
The goal of our analysis is trying to predict whether customers were overall satisfied or not of their flight experience.

## Variable description
* ...1: Index column
* Gender: Refers to the gender of the customer, either male or female.
* Customer Type: Indicates whether the customer is a loyal customer (someone who frequently chooses the same airline or company) or a disloyal customer (someone who doesn't consistently choose the same airline or company).
* Age: Represents the age of the customer.
* Type of Travel: Specifies the purpose of the customer's travel, such as personal travel or business travel.
* Customer Class: Indicates the class or category in which the customer traveled, such as Economy Plus or Business.
* Flight Distance: Represents the distance traveled by the customer's flight.
* Inflight Wi-Fi Service: Rates the satisfaction level of the customer with the inflight Wi-Fi service on a scale of 0 to 5.
* Departure/Arrival Time Convenience: Rates the satisfaction level of the customer with the convenience of departure and arrival times on a scale of 0 to 5.
* Ease of Online Booking: Rates the satisfaction level of the customer with the ease of booking flights online on a scale of 0 to 5.
* Gate Location: Rates the satisfaction level of the customer with the gate location at the airport on a scale of 0 to 5.
* Food and Drink: Rates the satisfaction level of the customer with the quality of food and drink provided on the flight on a scale of 0 to 5.
* Online Boarding: Rates the satisfaction level of the customer with the online boarding process on a scale of 0 to 5.
* Seat Comfort: Rates the satisfaction level of the customer with the comfort of the seats on the flight on a scale of 0 to 5.
* Inflight Entertainment: Rates the satisfaction level of the customer with the inflight entertainment options on a scale of 0 to 5.
* Onboard Service: Rates the satisfaction level of the customer with the service provided by the flight crew on a scale of 0 to 5.
* Leg Room Service: Rates the satisfaction level of the customer with the legroom space on the flight on a scale of 0 to 5.
* Baggage Handling: Rates the satisfaction level of the customer with the handling of baggage by the airline on a scale of 1 to 5.
* Check-in Service: Rates the satisfaction level of the customer with the check-in service at the airport on a scale of 0 to 5.
* Inflight Service: Rates the satisfaction level of the customer with the overall inflight service on a scale of 0 to 5.
* Cleanliness: Rates the satisfaction level of the customer with the cleanliness of the aircraft on a scale of 0 to 5.
* Departure Delay in Minutes: Specifies the number of minutes of departure delay experienced by the customer.
* Arrival Delay in Minutes: Specifies the number of minutes of arrival delay experienced by the customer.
* Satisfaction: Represents the overall satisfaction level of the customer with their travel experience, categorized as either "satisfied" or "neutral or dissatisfied"

# OUTLINE 
1. **Data upload**
2. **Exploratory analysis:**
    - missing values handling
    - outliers analysis
    - variables density
    - customer satisfaction distribution for categorical variables
    - customer satisfaction distribution for continuous variables
    - correlation matrix
3. **Train and test sets split**
4. **Logistic regression:**
    - general logistic model
    - forward step-wise models analysis
    - cross-validation
    - models comparison based on R-squared, CV error, CP
    - models selection (maybe one model per criteria if they differs)
    - models evaluation (accuracy, specificity, sensitivity, ...)
    - ROC curve
5. **Linear Discriminant Analysis:**
    - ...
    - compare LDA and logistic
6. **Tree predictors:**
    - ...
    - comparison with logistic model
7. **Conclusion and Interpretation**


## Libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(grid)
library(corrr)
```


# 1) DATA
```{r}
df <- read_csv("Database/airline_passenger_satisfaction.csv")
```

```{r}
head(df)
```

### Dataset structure
```{r}
str(df)
```


```{r}
summary(df)
# All the customer-generated variables are on a scales from 0 to 5 except for baggage handling that is on a scale from 1 to 5.
```




# 2) EXPLANATORY ANALYSIS

## Handling missing values
```{r}
# Function that counts NA values for each column
na_counts <- sapply(df, function(x) sum(is.na(x)))


# Calculate the percentage of NA values for each column
na_percentage <- na_counts / nrow(df) * 100

# Display a table with NA counts and percentages
na_table <- data.frame(Column = names(na_counts), NA_Count = na_counts, Percentage = na_percentage, row.names = NULL)

# na_table <- na_table[-1,]
na_table
```
 
```{r}
# Count NA values in departure delay when the arrival delay is zero
zero_count <- sum(df$departure_delay_in_minutes == 0 & is.na(df$arrival_delay_in_minutes))

zero_count
```
 
 
Missing values are present only for the column *arrival delay in minutes* for just a little bit more than 0.3% of the observations. We check if, as common sense might tell us, there is correlation between *departure delay in minutes.*
  
```{r}
correlation_dep_arr <- cor(df$departure_delay_in_minutes, df$arrival_delay_in_minutes, use = "complete.obs")

correlation_dep_arr
```
 
The correlation is quite high, we then assume reasonable populationg the *arrival delay in minutes* NA using the variable *departure delay in minutes.*
```{r}
df$arrival_delay_in_minutes <- ifelse(is.na(df$arrival_delay_in_minutes), df$departure_delay_in_minutes, df$arrival_delay_in_minutes)


sum(is.na(df$arrival_delay_in_minutes))
```
 
 
```{r}
# Alternative df where the NA are remove instead
df_na_deleted <- read_csv("Database/airline_passenger_satisfaction.csv")
df_na_deleted <- na.omit(df_na_deleted)
# df <- df_na_deleted
```
 
 
 
## Outliers analysis
The majority of variables take values on a fixed scale.
The variables worth checking for outliers are: *departure delay in minutes*, *arrival delay in minutes*, *flight distance*. 

```{r}
par(mfrow = c(1, 2))

plot(df$departure_delay_in_minutes, main = "Departure Delay",
     xlab = "Index", ylab = "Departure delay in minutes", col = "darkred")

plot(df$arrival_delay_in_minutes, main = "Arrival Delay",
     xlab = "Index", ylab = "Arrival delay in minutes", col = "darkgreen")

```
[...]
 
```{r}
Q1 <- quantile(df$departure_delay_in_minutes, 0.25)
Q3 <- quantile(df$departure_delay_in_minutes, 0.75)
IQR <- Q3 - Q1

lower_threshold <- Q1 - 1.5 * IQR
upper_threshold <- Q3 + 1.5 * IQR

iqr_outliers <- df$departure_delay_in_minutes[df$departure_delay_in_minutes < Q1 - 1.5 * IQR |
                                              df$departure_delay_in_minutes > Q3 + 1.5 * IQR]
iqr_outlier_count <- length(iqr_outliers)

c("Outliers identified with the IQR method:", iqr_outlier_count)
c("Percentage of outliers identified:", iqr_outlier_count / length(df$departure_delay_in_minutes) * 100)
```


```{r}
Q1 <- quantile(df$arrival_delay_in_minutes, 0.25)
Q3 <- quantile(df$arrival_delay_in_minutes, 0.75)
IQR <- Q3 - Q1

lower_threshold <- Q1 - 1.5 * IQR
upper_threshold <- Q3 + 1.5 * IQR

iqr_outliers <- df$arrival_delay_in_minutes[df$arrival_delay_in_minutes < Q1 - 1.5 * IQR |
                                              df$arrival_delay_in_minutes > Q3 + 1.5 * IQR]
iqr_outlier_count <- length(iqr_outliers)

c("Outliers identified with the IQR method:", iqr_outlier_count)
c("Percentage of outliers identified:", iqr_outlier_count / length(df$departure_delay_in_minutes) * 100)
```
The IQR method does not look suitable to identify outliers because it identify almost 14% of observations as outliers.
[For now I don't really know which method to use for identify what outliers to remove]





 
```{r}
plot(df$flight_distance, main = "Flight Distance",
     xlab = "Index", ylab = "Flight distance", col = "darkblue")
```
[...]


```{r}
# plot(df$age, main = "Customers' Age",
#     xlab = "Index", ylab = "Age")
# Is it worth plotting age too?
```


## Data cleaning
[I'm gonna remove the ...1 column because it is just an extra index column]


































