---
title: "AIRLINE PASSENGERS SATISFACTION: CLASSIFICATION"
output: html_notebook
date: 16/06/2023
---

## Database description
The following database contains information about airline customers, the flight they took and the results of a satisfaction questionnaire about their flight experience.
The goal of our analysis is trying to predict whether customers were overall satisfied or not of their flight experience.

## Variable description
* ...1: Index column
* Gender: Refers to the gender of the customer, either male or female.
* Customer Type: Indicates whether the customer is a loyal customer (someone who frequently chooses the same airline or company) or a disloyal customer (someone who doesn't consistently choose the same airline or company).
* Age: Represents the age of the customer.
* Type of Travel: Specifies the purpose of the customer's travel, such as personal travel or business travel.
* Customer Class: Indicates the class or category in which the customer traveled, such as Economy Plus or Business.
* Flight Distance: Represents the distance traveled by the customer's flight.
* Inflight Wi-Fi Service: Rates the satisfaction level of the customer with the inflight Wi-Fi service on a scale of 0 to 5.
* Departure/Arrival Time Convenience: Rates the satisfaction level of the customer with the convenience of departure and arrival times on a scale of 0 to 5.
* Ease of Online Booking: Rates the satisfaction level of the customer with the ease of booking flights online on a scale of 0 to 5.
* Gate Location: Rates the satisfaction level of the customer with the gate location at the airport on a scale of 0 to 5.
* Food and Drink: Rates the satisfaction level of the customer with the quality of food and drink provided on the flight on a scale of 0 to 5.
* Online Boarding: Rates the satisfaction level of the customer with the online boarding process on a scale of 0 to 5.
* Seat Comfort: Rates the satisfaction level of the customer with the comfort of the seats on the flight on a scale of 0 to 5.
* Inflight Entertainment: Rates the satisfaction level of the customer with the inflight entertainment options on a scale of 0 to 5.
* Onboard Service: Rates the satisfaction level of the customer with the service provided by the flight crew on a scale of 0 to 5.
* Leg Room Service: Rates the satisfaction level of the customer with the legroom space on the flight on a scale of 0 to 5.
* Baggage Handling: Rates the satisfaction level of the customer with the handling of baggage by the airline on a scale of 1 to 5.
* Check-in Service: Rates the satisfaction level of the customer with the check-in service at the airport on a scale of 0 to 5.
* Inflight Service: Rates the satisfaction level of the customer with the overall inflight service on a scale of 0 to 5.
* Cleanliness: Rates the satisfaction level of the customer with the cleanliness of the aircraft on a scale of 0 to 5.
* Departure Delay in Minutes: Specifies the number of minutes of departure delay experienced by the customer.
* Arrival Delay in Minutes: Specifies the number of minutes of arrival delay experienced by the customer.
* Satisfaction: Represents the overall satisfaction level of the customer with their travel experience, categorized as either "satisfied" or "neutral or dissatisfied"

# OUTLINE 
1. **Data upload**
2. **Exploratory analysis:**
    - missing values handling
    - outliers analysis:
        + IQR method
        + 3*SD method
    - data cleaning
    - descriptive analysis:
        + satisfaction distribution
        + customers information
        + customer-generated variables
        + numerical variables
    - correlation matrix
3. **Train and test sets split**
4. **Logistic regression:**
    - general logistic model
    - forward step-wise models analysis
    - cross-validation
    - models comparison based on R-squared, CV error, CP
    - models selection (maybe one model per criteria if they differs)
    - models evaluation (accuracy, specificity, sensitivity, ...)
    - ROC curve
5. **Linear Discriminant Analysis:**
    - ...
    - compare LDA and logistic
6. **Tree predictors:**
    - ...
    - comparison with logistic model
7. **Conclusion and Interpretation**


## Libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)
library(corrr)
library(ggradar)
library(scales)
library(tibble)
# ggradar needs to be installed from github 
# > install.packages("remotes")
# > remotes::install_github("ricardo-bion/ggradar")
```


# **1) DATA**
```{r}
df <- read_csv("Database/airline_passenger_satisfaction.csv")
```

```{r}
head(df)
```

### Dataset structure
```{r}
str(df)
```


```{r}
summary(df)
# All the customer-generated variables are on a scales from 0 to 5 except for baggage handling that is on a scale from 1 to 5.
```




# **2) EXPLANATORY ANALYSIS**

## Handling missing values
```{r}
# Function that counts NA values for each column
na_counts <- sapply(df, function(x) sum(is.na(x)))


# Calculate the percentage of NA values for each column
na_percentage <- na_counts / nrow(df) * 100

# Display a table with NA counts and percentages
na_table <- data.frame(Column = names(na_counts), NA_Count = na_counts, Percentage = na_percentage, row.names = NULL)

# na_table <- na_table[-1,]
na_table
```
 
```{r}
# Count NA values in departure delay when the arrival delay is zero
zero_count <- sum(df$departure_delay_in_minutes == 0 & is.na(df$arrival_delay_in_minutes))

zero_count
```
 
 
Missing values are present only for the column *arrival delay in minutes* for just a little bit more than 0.3% of the observations. We check if, as common sense might tell us, there is correlation between *departure delay in minutes.*
  
```{r}
correlation_dep_arr <- cor(df$departure_delay_in_minutes, df$arrival_delay_in_minutes, use = "complete.obs")

correlation_dep_arr
```
 
The correlation is quite high, we then assume reasonable populationg the *arrival delay in minutes* NA using the variable *departure delay in minutes.*
```{r}
df$arrival_delay_in_minutes <- ifelse(is.na(df$arrival_delay_in_minutes), df$departure_delay_in_minutes, df$arrival_delay_in_minutes)


sum(is.na(df$arrival_delay_in_minutes))
```
 
 
```{r}
# Alternative df where the NA are remove instead
df_na_deleted <- read_csv("Database/airline_passenger_satisfaction.csv")
df_na_deleted <- na.omit(df_na_deleted)
# df <- df_na_deleted
```
 
 
 
## Outliers analysis
The majority of variables take values on a fixed scale.
The variables worth checking for outliers are: *departure delay in minutes*, *arrival delay in minutes*, *flight distance*. 

```{r}
par(mfrow = c(1, 2))

plot(df$departure_delay_in_minutes, main = "Departure Delay",
     xlab = "Index", ylab = "Departure delay in minutes", col = "darkred")

plot(df$arrival_delay_in_minutes, main = "Arrival Delay",
     xlab = "Index", ylab = "Arrival delay in minutes", col = "darkgreen")

```
[...]


## Inter Quantile Range method 
```{r}
Q1 <- quantile(df$departure_delay_in_minutes, 0.25)
Q3 <- quantile(df$departure_delay_in_minutes, 0.75)
IQR <- Q3 - Q1

lower_threshold <- Q1 - 1.5 * IQR
upper_threshold <- Q3 + 1.5 * IQR

iqr_outliers <- df$departure_delay_in_minutes[df$departure_delay_in_minutes < Q1 - 1.5 * IQR |
                                              df$departure_delay_in_minutes > Q3 + 1.5 * IQR]
iqr_outlier_count <- length(iqr_outliers)

c("Outliers identified with the IQR method:", iqr_outlier_count)
c("Percentage of outliers identified:", iqr_outlier_count / length(df$departure_delay_in_minutes) * 100)
```


```{r}
Q1 <- quantile(df$arrival_delay_in_minutes, 0.25)
Q3 <- quantile(df$arrival_delay_in_minutes, 0.75)
IQR <- Q3 - Q1

lower_threshold <- Q1 - 1.5 * IQR
upper_threshold <- Q3 + 1.5 * IQR

iqr_outliers <- df$arrival_delay_in_minutes[df$arrival_delay_in_minutes < Q1 - 1.5 * IQR |
                                              df$arrival_delay_in_minutes > Q3 + 1.5 * IQR]
iqr_outlier_count <- length(iqr_outliers)

c("Outliers identified with the IQR method:", iqr_outlier_count)
c("Percentage of outliers identified:", iqr_outlier_count / length(df$departure_delay_in_minutes) * 100)
```
The IQR method does not look suitable to identify outliers because it identify almost 14% of observations as outliers.
[For now I don't really know which method to use for identify what outliers to remove]


## SD method
```{r}
z_scores_dep <- scale(df$departure_delay_in_minutes)
z_scores_arr <- scale(df$arrival_delay_in_minutes)

outliers_dep <- which(abs(z_scores_dep) > 3)
outliers_arr <- which(abs(z_scores_arr) > 3)

sd_outliers_count_dep <- length(outliers_dep)
sd_outliers_count_arr <- length(outliers_arr)

c("Observations that falls over 3 times the sd (departure delay):", sd_outliers_count_dep)
c("Percentage of outliers identified (departure delay):", sd_outliers_count_dep / length(df$departure_delay_in_minutes) * 100)

c("Observations that falls over 3 times the sd (arrival delay):", sd_outliers_count_arr)
c("Percentage of outliers identified (arrival delay):", sd_outliers_count_arr / length(df$arrival_delay_in_minutes) * 100)
```
Over 2% of the observations are identified as outliers using the standard deviation method

 
```{r}
plot(df$flight_distance, main = "Flight Distance",
     xlab = "Index", ylab = "Flight distance", col = "darkblue")
```
[...]

```{r}
z_scores_dist <- scale(df$flight_distance)

outliers_dist <- which(abs(z_scores_dist) > 3)

sd_outliers_count_dist <- length(outliers_dist)

c("Observations that falls over 3 times the sd (flight distance):", sd_outliers_count_dist)
c("Percentage of outliers identified (flight distance):", sd_outliers_count_dist / length(df$flight_distance) * 100)
```

[For now I will not remove any outliers]

```{r}
# plot(df$age, main = "Customers' Age",
#     xlab = "Index", ylab = "Age")
# Is it worth plotting age too?
```



## Data cleaning
Deleting the *...1* column because it is just an extra index column
```{r}
df <- df[, -1]
# df
```
Renaming Gender variable to gender for consistency
```{r}
df <- df %>% 
  rename(gender = Gender)
```

 
 
 
# Descriptive analysis

```{r}
p_satisf <- ggplot(df, aes(x = satisfaction)) +
  geom_bar(aes(fill = satisfaction)) +
  geom_text(aes(y = ..count.., 
                label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(0.3), 
            size = 4,
            vjust = - 0.5) +
  scale_fill_manual(values = c("red", "green")) +
  labs(title = "Satisfaction Distribution", y = "Count")

p_satisf
```

### Define function to plot bar charts
```{r}
# Plot bar chart function
plot_bar_chart <- function(df, var_name) {
  var_name <- enquo(var_name)  # Defuse the var_name so it can be evaluated later

  plot_var <- ggplot(df, aes(x = !!var_name)) +  # !! to inject the var_name back to evaluate it
    geom_bar(aes(fill = satisfaction)) +
    geom_text(aes(y = ..count.., 
                  label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
              stat = 'count', 
              position = position_dodge(0.3), 
              size = 4,
              vjust = -0.5) +
    scale_fill_manual(values = c("red", "green")) +
    labs(title = paste0("Satisfaction Distribution for ", quo_name(var_name)), 
         x = quo_name(var_name), y = "Count")  # quo_name() to return the var_name as string

  print(plot_var)
}
```


### Customers information Variables
```{r}
# Gender
plot_bar_chart(df, gender)

# Age
df$age_group <- cut(df$age, breaks = seq(0, max(df$age) + 10, by = 10), right = FALSE)
plot_bar_chart(df, age_group)

# Customer type
plot_bar_chart(df, customer_type)

# Type of travel
plot_bar_chart(df, type_of_travel)

# Customer class
plot_bar_chart(df, customer_class)
```


### Customer-generated variables
```{r}
# Inflight Wifi service
plot_bar_chart(df, inflight_wifi_service)

# Departure/arrival time convenient
plot_bar_chart(df, departure_arrival_time_convenient)

# Ease of online booking
plot_bar_chart(df, ease_of_online_booking)

# Gate location
plot_bar_chart(df, gate_location)

# Food and drink
plot_bar_chart(df, food_and_drink)

# Online boarding
plot_bar_chart(df, online_boarding)

# Seat comfort
plot_bar_chart(df, seat_comfort)

# Inflight entertainment
plot_bar_chart(df, inflight_entertainment)

# Onboard service
plot_bar_chart(df, onboard_service)

# Leg room service
plot_bar_chart(df, leg_room_service)

# Baggage handling
plot_bar_chart(df, baggage_handling)

# Check-in service
plot_bar_chart(df, checkin_service)

# Inflight service
plot_bar_chart(df, inflight_service)

# Cleanliness
plot_bar_chart(df, cleanliness)

```


### Numerical Variables
```{r}
# Flight distance
p_distance <- ggplot(data = df, aes(flight_distance, color = satisfaction)) +
  geom_freqpoly(binwidth = 100, linewidth = 1) +
  scale_color_manual(values = c("red", "green")) +
  labs(title = "Satisfaction distribution for Flight distance")

# Distance density
p_distancr_den <- ggplot(df, aes(x = flight_distance, fill = satisfaction)) +
  geom_density(alpha = 0.5) + 
  labs(title = "Density Plot of Flight distance", 
       x = "flight_distance", y = "Density") + 
  scale_fill_manual(values = c("red", "green"), name = "Satisfaction")

grid.arrange(p_distance, p_distancr_den)

# Departure delay
# I selected delays under 200 min otherwise the graphs look even worse than these
p_dep_del <- ggplot(data = df[df$departure_delay_in_minutes < 200, ], 
                    aes(departure_delay_in_minutes, color = satisfaction)) +
  geom_freqpoly(binwidth = 15, linewidth = 0.7) +
  scale_color_manual(values = c("red", "green")) +
  labs(title = "Satisfaction distribution for Departure delay")
p_dep_del

# Arrival delay
p_arr_del <- ggplot(data = df[df$arrival_delay_in_minutes < 200, ], 
                    aes(arrival_delay_in_minutes, color = satisfaction)) +
  geom_freqpoly(binwidth = 15, linewidth = 0.7) +
  scale_color_manual(values = c("red", "green")) +
  labs(title = "Satisfaction distribution for Arrival distance")
p_arr_del

```
 


# Correlation matrix for the customer-generated variables
```{r}
correlation_matrix <- df %>% 
  select(inflight_wifi_service,departure_arrival_time_convenient, 
         ease_of_online_booking, gate_location, food_and_drink, online_boarding,
         seat_comfort, inflight_entertainment, onboard_service, leg_room_service,
         baggage_handling, checkin_service, inflight_service, cleanliness) %>% 
  cor()

round(correlation_matrix, 3)
```

### Heatmap
```{r}
heatmap(correlation_matrix, 
        col = colorRampPalette(c("blue", "white", "red"))(1000),
        main = "Correlation Heatmap", Colv = NA, margins = c(15, 9))

```




## Radar map
```{r}
exclude_cols <- c("gender", "customer_type", "age", "type_of_travel", 
                  "customer_class", "flight_distance", 
                  "departure_delay_in_minutes", "arrival_delay_in_minutes") 
                  

df_mean <- df %>%
  select(-exclude_cols) %>%
  group_by(satisfaction) %>%
  summarize(across(any_of(names(.)[sapply(., is.numeric) | sapply(., is.logical)]), 
                   ~ mean(., na.rm = TRUE)))
df_mean

```
```{r}
# df_mean[, 2:15]

radarchart(rbind(rep(5, 14), rep(0, 14), df_mean[, 2:15]),
           vlabels = colnames(df_mean[, 2:15]))
```

```{r}
ggradar(df_mean, grid.min = 0, grid.max = 5, 
        axis.label.size = 3,
        group.point.size = 3,
        group.line.width = 1) +
  theme(legend.position = c(1, 0),
        legend.spacing = unit(c(0, 0, 0, 0), "lines"),
        legend.key.size = unit(c(0, 0, 0, 0), "lines"),
        legend.text = element_text(size = 7),
        plot.margin = unit(c(2, 8, 2, 8), "lines")) +
  scale_color_manual(values = c("red", "green"))
```











